---
title: "LK of Excursion sets"
author: "Zinedine Bounnah"
date: "2023-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plotly)
library(RandomFields)
library(raster)
library(RColorBrewer)
library(plotly)
library(stars)
library(ggplot2)
library(cowplot)

```

Simulation of a Gaussian random field :

```{r}
## define the locations:
from <- 0
to <- 2^8
x.seq <- seq(from, to, by=0.75) 
y.seq <- seq(from, to, by=0.75)


# Simulate the random field
rf=RFsimulate(RMexp(scale=sqrt(2/ka)), x=x.seq,y=y.seq)


ka=(100/2^8)^2
x.seq <- seq(from, to, by=0.75)
y.seq <- seq(from, to, by=0.75)
rf=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)

# white/black color palette 
pal <- colorRampPalette(c("white","black"))
plot(raster::raster(rf),col=pal(7))

```

3D plot of the Gaussian Field

```{r}
##plot 3D
df <- expand.grid(x = x.seq, y = y.seq)
df$z <- rf@data[["variable1"]]

plot=plot_ly() %>% 
  add_trace(data = df,  x=df$x, y=df$y, z=df$z, type="mesh3d" ) %>%
  layout(title = "Gaussian Field with covariance exp(-r²)") 

plot
```

Chi 2 :

```{r}


x.seq <- seq(from, to, by=0.75)
y.seq <- seq(from, to, by=0.75)
  
  
# Simulate the random field
rf=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf2=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)

rf@data[["variable1"]]=rf@data[["variable1"]]^2+rf2@data[["variable1"]]^2
rf@data[["variable1"]]=0.5*(rf@data[["variable1"]]-2)


# Define color palette for plotting
pal <- colorRampPalette(c("white", "black"))

# Convert simulated random field to a raster object and plot it
r <- raster(rf)

plot(r, col = pal(7))


```

t field :

```{r}
k=4


  
x.seq <- seq(from, to, by=0.75) 
y.seq <- seq(from, to, by=0.75)

  
rf=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf1=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf2=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf3=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf4=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)

chi=rf1@data[["variable1"]]^2+rf2@data[["variable1"]]^2+rf3@data[["variable1"]]^2+rf4@data[["variable1"]]^2
rf@data[["variable1"]]=(rf@data[["variable1"]])/sqrt(chi/4)
rf@data[["variable1"]]=rf@data[["variable1"]]*sqrt(1/2)


# Define color palette for plotting
pal <- colorRampPalette(c("white", "black"))

# Convert simulated random field to a raster object and plot it
r <- raster(rf)
plot(r, col = pal(50))

```


Function to get the excursion set at level u of a random field:

```{r}
Excursion_set=function(randomfield,level){
  # Set the threshold
  excursion_mask <- randomfield@data >= level
  
  # Create a copy of randomfield 
  excursionset <- randomfield
  
  # Set values outside the excursion set to NA
  excursionset@data[!excursion_mask] <- NA
  
  return(excursionset)
}
```

Example and plots :

```{r}
## define the locations:
from <- 0
to <- 2^8
x.seq <- seq(from, to, by=0.75) 
y.seq <- seq(from, to, by=0.75) 


# Simulate the random field
rf=RFsimulate(RMexp(scale=sqrt(7)/ka), x=x.seq,y=y.seq)

plot(raster(rf),col=pal(7))

##PLOT WITHIN T 
level=0
# Convert it to a raster and plot
excursionset=Excursion_set(rf,level)
exc_raster=raster(excursionset)
plot(exc_raster>=level,col=pal(2),legend=FALSE)


##PLOT WITHIN T 
level=1
# Convert it to a raster and plot
excursionset=Excursion_set(rf,level)
exc_raster=raster(excursionset)
plot(exc_raster>=level,col=pal(2),legend=FALSE)



##PLOT WITHIN T 
level=2
# Convert it to a raster and plot
excursionset=Excursion_set(rf,level)
exc_raster=raster(excursionset)
plot(exc_raster>=level,col=pal(2),ylim=c(0,2^8),legend=FALSE)





##3D plot
df <- expand.grid(x = x.seq, y = y.seq)
df$z <- excursionset@data[["variable1"]]

plot3d <- plot_ly(data = df, x = ~x, y = ~y, z = ~z, type = "mesh3d") %>%
  layout(title = "Excursion set at level u=0",scene=list(zaxis = list(range = c(-4,4))))

plot3d


```



```{r}


Euler_charac2=function(randomfield,level){ ##randomfield class "RFspatialGridDataFrame"
  
  # Compute the excursion set
  excursionset <- Excursion_set(randomfield,level)
  excursionsett=st_as_stars(excursionset) >= level
  
  exc_sf <- st_as_sf(excursionsett, merge = TRUE)
  connected_comp=nrow(exc_sf)
  
  # Set the threshold
  inv_mask <- randomfield@data < level
  # Create copy of the randomfield 
  inv_exc <- randomfield
  # Set values outside the complementary of the excursion set to NA
  inv_exc@data[!inv_mask] <- NA
  
  invexc=st_as_stars(inv_exc) >= level
  inv_sf <- st_as_sf(invexc, merge = TRUE)
  holes=nrow(inv_sf)-1
  
  euler_characteristic=connected_comp - holes
  
  
  return(list(eul=euler_characteristic, exc=excursionset,invexc=inv_exc))
}
```



Area of the excursion set : Proportion of colored pixels \* area of the window We do it this way to allow a discretization of the interval that is not only xi=1,2..n

```{r}
Area_exc2=function(excursionset,level){ ##excursionset must be of class "RFspatialGridDataFrame", level is the level for the excursion set
  
  # Extract window size 
  maxvalues=excursionset@grid@cellsize * (excursionset@grid@cells.dim-1) ##length of the axis
  minvalues=excursionset@grid@cellcentre.offset ##where the window start
  
  # Total area
  totarea=prod(maxvalues-minvalues) 
  
  
  # Convert excursionset to a binary raster
  excursionset=raster(excursionset)
  excursionset=excursionset >= level
  
  # Convert logical to 0s and 1s 
  binary_image <- as.integer(excursionset[] == TRUE)
  
  # Get the area by looking at the percentage of colored pixels in the image
  proportion=sum(binary_image == 1, na.rm = TRUE) / length(binary_image) 
  
  area=proportion*totarea
  
  return(area)
}
```

Perimeter

Example of the process is shown later

```{r}
Perim_exc=function(excursionset,level){ ##randomfield must be of class "RFspatialGridDataFrame", level is the level for the excursion set
  ## Convert excursionset to a binary stars
  excursionsett=st_as_stars(excursionset)>=level
  
  ## Now convert to sf class
  exc_sf=st_as_sf(excursionsett,merge=TRUE)
  
  ## Compute the perimeter using st_length function. we use st_cast() with parameter "MULTILINESTRING" for the function to work properly
  perimeter=sum(st_length(st_cast(exc_sf,"MULTILINESTRING")))
  
  
  return(perimeter)
}
```


```{r}
from <- 0
to <- 2^8
ka <- (100/2^8)^2
x.seq <- seq(from, to, by=1)
y.seq <- seq(from, to, by=1)
rf <- RFsimulate(RMexp(scale=sqrt(2)/ka), x = x.seq, y = y.seq)


level <- 1
excursionset <- Excursion_set(rf, level)
excursionsett <- st_as_stars(excursionset) >= level

exc_sf <- st_as_sf(excursionsett, merge = TRUE)
merged <- st_cast(exc_sf, "MULTILINESTRING")

plot(excursionsett, main = NULL, breaks = "equal")
graphics::plot(exc_sf, main = NULL)
graphics::plot(merged, main = NULL)

```

All lipschitz killing together :

```{r}
LipschitzKilling=function(randomfield,level){
  excursionset=Excursion_set(randomfield,level)
  L0=Euler_charac2(randomfield,level)
  L1=Perim_exc(excursionset,level)/2
  L2=Area_exc2(excursionset,level)
  return(list(L0=L0$eul, L1=L1,L2=L2))
}
```


Monte Carlo

```{r}
MCLK=function(n,from,to,level,model,ka=1) {

  
L0s=numeric(n)
L1s=numeric(n)
L2s=numeric(n)

for(i in 1:n){
    ## define the locations:



# Simulate the random field
if(model=="Gaussian"){
x.seq <- seq(from, to, by=1)
y.seq <- seq(from, to, by=1)
  rf=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
}
if(model=="chi"){
x.seq <- seq(from, to, by=1)
y.seq <- seq(from, to, by=1)
  
  
# Simulate the random field
rf=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf2=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)

rf@data[["variable1"]]=rf@data[["variable1"]]^2+rf2@data[["variable1"]]^2
rf@data[["variable1"]]=0.5*(rf@data[["variable1"]]-2)
}
if(model=='t'){
  
x.seq <- seq(from, to, by=1) 
y.seq <- seq(from, to, by=1)

  
rf=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf1=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf2=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf3=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)
rf4=RFsimulate(RMexp(scale=sqrt(2)/ka), x=x.seq,y=y.seq)

chi=rf1@data[["variable1"]]^2+rf2@data[["variable1"]]^2+rf3@data[["variable1"]]^2+rf4@data[["variable1"]]^2
rf@data[["variable1"]]=(rf@data[["variable1"]])/sqrt(chi/4)
rf@data[["variable1"]]=rf@data[["variable1"]]*sqrt(1/2)
}

LK=LipschitzKilling(rf,level)
L0s[i]=LK$L0[1]
L1s[i]=LK$L1[1]
L2s[i]=LK$L2[1]
}

L0=mean(L0s)
L0CI=list(-1.96*sd(L0s)/sqrt(n),1.96*sd(L0s)/sqrt(n))
L1=mean(L1s)
L1CI=list(-1.96*sd(L1s)/sqrt(n),1.96*sd(L1s)/sqrt(n))
L2=mean(L2s)
L2CI=list(-1.96*sd(L2s)/sqrt(n),1.96*sd(L2s)/sqrt(n))
return(list(L0,L1,L2,L0CI,L1CI,L2CI))
}

LKD=function(MC,lk1,lk2,lk3){
L0S=numeric(length(MC))
L1S=numeric(length(MC))
L2S=numeric(length(MC))
lowery=numeric(length(MC))
uppery=numeric(length(MC))
lowerz=numeric(length(MC))
upperz=numeric(length(MC))
lowert=numeric(length(MC))
uppert=numeric(length(MC))
for(i in 1:length(MC)){
  L0S[i]=MC[[i]][[1]]/(to*to)
  L1S[i]=MC[[i]][[2]]/(to*to)
  L2S[i]=MC[[i]][[3]]/(to*to)
  lowery[i]=MC[[i]][[4]][[1]]/(to*to)
  uppery[i]=MC[[i]][[4]][[2]]/(to*to)
  lowerz[i]=MC[[i]][[5]][[1]]/(to*to)
  upperz[i]=MC[[i]][[5]][[2]]/(to*to)
  lowert[i]=MC[[i]][[6]][[1]]/(to*to)
  uppert[i]=MC[[i]][[6]][[2]]/(to*to)
}

data=data.frame(u=levels,L0=L0S,L1=L1S,L2=L2S,lowery=lowery,uppery=uppery,lowerz=lowerz,upperz=upperz,lowert=lowert,uppert=uppert)


par(mfrow=c(2,3), oma=c(0,0,2,0));

p1=ggplot(data, aes(u, L0)) +
 geom_point(shape=1) +
  geom_errorbar(
    aes(ymin = L0S+lowery, ymax = L0S+uppery),
    width = 0.3,colour="red") +
    geom_function(fun = lk1)

p2=ggplot(data, aes(u, L1)) +
 geom_point(shape=1) +
  geom_errorbar(
    aes(ymin = L1S+lowerz, ymax = L1S+upperz),
    width = 0.3,colour="red") +
    geom_function(fun = lk2)

p3=ggplot(data, aes(u, L2)) +
 geom_point(shape=1) +
  geom_errorbar(
    aes(ymin = L2S+lowert, ymax = L2S+uppert),
    width = 0.3,colour="red") +
    geom_function(fun = lk3)



##unbiased

L0s=numeric(length(MC))
L1s=numeric(length(MC))
L2s=numeric(length(MC))
lowery=numeric(length(MC))
uppery=numeric(length(MC))
lowerz=numeric(length(MC))
upperz=numeric(length(MC))
lowert=numeric(length(MC))
uppert=numeric(length(MC))
for(i in 1:length(MC)){
  L0s[i]=MC[[i]][[1]]/(to*to)-(to*4)*(MC[[i]][[2]]/(to*to))/(pi*to*to)+((to*4)^2/(2*pi*(to*to)^2) -1/(to*to) )*(MC[[i]][[3]]/(to*to))
  L1s[i]=MC[[i]][[2]]/(to*to)-(to*4)*(MC[[i]][[3]]/(to*to))/(2*to*to)
  L2s[i]=MC[[i]][[3]]/(to*to)
  lowery[i]=MC[[i]][[4]][[1]]/(to*to)
  uppery[i]=MC[[i]][[4]][[2]]/(to*to)
  lowerz[i]=MC[[i]][[5]][[1]]/(to*to)
  upperz[i]=MC[[i]][[5]][[2]]/(to*to)
  lowert[i]=MC[[i]][[6]][[1]]/(to*to)
  uppert[i]=MC[[i]][[6]][[2]]/(to*to)
  
}
data=data.frame(u=levels,L0=L0s,L1=L1s,L2=L2s,lowery=lowery,uppery=uppery,lowerz=lowerz,upperz=upperz,lowert=lowert,uppert=uppert)

par(mfrow=c(2,3), oma=c(0,0,2,0));

p4=ggplot(data, aes(u, L0)) +
 geom_point(shape=1) +
  geom_errorbar(
    aes(ymin = L0s+lowery, ymax = L0s+uppery),
    width = 0.3,
    position=position_dodge(width=.2),colour="red") +
    geom_function(fun = lk1)

p5=ggplot(data, aes(u, L1)) +
 geom_point(shape=1) +
  geom_errorbar(
    aes(ymin = L1s+lowerz, ymax = L1s+upperz),
    width = 0.3,
    position=position_dodge(width=.2),colour="red") +
    geom_function(fun = lk2)

p6=ggplot(data, aes(u, L2)) +
 geom_point(shape=1) +
  geom_errorbar(
    aes(ymin = L2s+lowert, ymax = L2s+uppert),
    width = 0.3,
    position=position_dodge(width=.2),colour="red") +
    geom_function(fun = lk3)


plot_grid(p1,p2,p3,p4,p5,p6)
  
}


```


Gaussian exp(-ka r²)
```{r}
n=2
ka=(100/2^8)^2
from=0
to=2^8
levels=seq(-5,5,by=0.5)

lk1=function(u){
  (2*pi)^(-3/2) * 2 * ka * u * exp(-u^2/2)
}
lk2=function(u){
  (0.25*sqrt(2 * ka)*exp(-u^2/2))
}
lk3=function(u){
  (1-pnorm(u))
}

MC=apply(as.matrix(levels),1,MCLK,from=from,to=to,n=n,model="Gaussian",ka=ka) #n=1 just for testing since the computation for a large n can take quite some time


LKD(MC,lk1,lk2,lk3)

```


Chi
exp -ka
```{r}
ka=(100/2^8)^2
from=0
to=2^8
levels=seq(-0.5,5,by=0.5)
levels=c(-0.75,levels) 
MC=apply(as.matrix(levels),1,MCLK,from=from,to=to,n=10,model="chi",ka=ka) #n=5 just for testing since the computation for a large n can take quite some time

lk1=function(u){
 2*ka*(u*2+1)*exp(-(u*2+2)/2)/(pi*2*gamma(1))
}
lk2=function(u){
  sqrt(pi*2*ka)*(2+u*2)^(0.5)*exp(-(2+u*2)/2)/(2^(3/2)*gamma(1))
}
lk3=function(u){
  1-pchisq(2+u*2,df=2)
}

LKD(MC,lk1,lk2,lk3)

```


t
```{r}
ka=(100/2^7)^2
from=0
to=2^7
levels=seq(-15,15,by=1)
start_time = Sys.time()
MC=apply(as.matrix(levels),1,MCLK,from=from,to=to,n=2,model="t",ka=ka) #n=5 just for testing since the computation for a large n can take quite some time

lk1=function(u){
 (6*ka)/(4*pi^(3/2))*u/sqrt(2)*gamma(3/2)/gamma(2)*(1+u^2/2)^(-3/2)
}
lk2=function(u){
  sqrt(2*ka)/4*(1+u^2/2)^(-3/2)
}
lk3=function(u){
  1-pt(u*sqrt(3),df=4)
}

LKD(MC,lk1,lk2,lk3)


```





